---
title: "Lab 4"
author: "Mikaela Blount"
date: "10/17/2019"
output: html_document
---

## Part two: Tidying (A)

1.
```{r}
defClean2 = defClean %>%
  gather(key = "Statistic", value = "Value",Tackles_Solo, Tackles_ASST, Tackles_TFL, Tackles_Sack, Turnover_FF, Turnover_FR, Turnover_INT, Pass_QBH, Pass_PB)
```

2. 
The rarest defensive skill is creating turnovers, espically interceptions (Turnover_INT). This makes sense because the big 12 is known for terrible defensives and high-power offenses; University of Oklahoma's last two quarterbacks were Heisman Winners. Iowa State also struggled for many years with talent on defensive until Matt Campell's second year.
```{r}
ggplot(defClean2, aes(x=Value)) +
    geom_histogram(position="identity", colour="grey40", bins = 10000) +
    facet_grid(. ~ Statistic)

```



3.
ISU had better defense against Kansas State because the points are higher on the y-axis than further on the x-axis indicating that ISU had more solo tackles against Kansas State. The majority of the players are at (0,0) because they recorded no sacks against either team.
```{r}
defKSWV = defClean2 %>%
  filter(Statistic == "Tackles_Solo" & Opponent_Opponent %in% c("West Virginia","Kansas State")) %>%
  spread(key = Statistic, value = Value)

d = defKSWV %>%
  spread(key = Opponent_Opponent,value = Tackles_Solo)

ggplot(d ,aes(y = `Kansas State`, x = `West Virginia`,color=Name)) + geom_point(na.rm = TRUE)
```



## Part two: Tidying  (B)
Now let's turn our attention to the player biography `bioClean`.

1. 
```{r}
bioClean2 = bioClean %>%
  separate(Hometown, c("City","State"), sep = ", ", remove = TRUE)

head(bioClean2)
```

2. 
```{r}
table(bioClean2$State)
```


