---
title: "Lab 4"
author: "Mikaela Blount and Jaden Birkner"
date: "10/24/2019"
output: html_document
---

#### Owner's GitHub username: mblount2
#### Partner's GitHub username: jebirkner
#### Repo name: ds202_lab4
#### Due date: Thursday Oct. 24 by 10 pm.

## Part one: Cleaning data

```{r}
library(tidyverse)
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

1.
```{r}
defense$Name = factor(defense$Name)
defense$Opponent_Opponent = factor(defense$Opponent_Opponent)
offense$Name = factor(offense$Name)
offense$Opponent_Opponent = factor(offense$Opponent_Opponent)
bio$Name = factor(bio$Name)
```

2.
```{r}
defClean = mutate_at(defense, , .funs = as.numeric, .vars = c("Tackles_Solo", "Tackles_ASST","Tackles_TFL", "Tackles_Sack", "Turnover_FF", "Turnover_FR", "Turnover_INT", "Pass_QBH", "Pass_PB"))
```

```{r}
offClean = mutate_at(offense, .funs = as.numeric, .vars = names(offense[c(3:12)]))
```

```{r}
bio = mutate_at(bio, .funs = as.numeric, .vars = c("Weight"))
```

3.
```{r}
bioClean = bio %>%
  separate(Height, into = c("Feet", "Inches"))
bioClean$Feet = as.numeric(bioClean$Feet)
bioClean$Inches = as.numeric(bioClean$Inches)
bioClean$Feet = bioClean$Feet * 12
bioClean = mutate(bioClean, Height = bioClean$Feet + bioClean$Inches)
bioClean = select(bioClean, "Name", "Position", "Weight", "Class",  "Hometown", "Highschool", "Height")
names(bioClean)
```

4.
```{r}
(str(defClean))
```

```{r}
(str(offClean))
```

```{r}
(str(bioClean))
```

## Part two: Tidying (A)

1.
```{r}
defClean2 = defClean %>%
  gather(key = "Statistic", value = "Value",Tackles_Solo, Tackles_ASST, Tackles_TFL, Tackles_Sack, Turnover_FF, Turnover_FR, Turnover_INT, Pass_QBH, Pass_PB)
```

2. 
The rarest defensive skill is creating turnovers, espically interceptions (Turnover_INT). This makes sense because the big 12 is known for terrible defensives and high-power offenses; University of Oklahoma's last two quarterbacks were Heisman Winners. Iowa State also struggled for many years with talent on defensive until Matt Campell's second year.
```{r}
ggplot(defClean2, aes(x=Value)) +
    geom_histogram(position="identity", colour="grey40", bins = 10000) +
    facet_grid(. ~ Statistic)

```



3.
ISU had better defense against Kansas State because the points are higher on the y-axis than further on the x-axis indicating that ISU had more solo tackles against Kansas State. The majority of the players are at (0,0) because they recorded no sacks against either team.
```{r}
defKSWV = defClean2 %>%
  filter(Statistic == "Tackles_Solo" & Opponent_Opponent %in% c("West Virginia","Kansas State")) %>%
  spread(key = Statistic, value = Value)

d = defKSWV %>%
  spread(key = Opponent_Opponent,value = Tackles_Solo)

ggplot(d ,aes(y = `Kansas State`, x = `West Virginia`,color=Name)) + geom_point(na.rm = TRUE)
```



## Part two: Tidying  (B)
Now let's turn our attention to the player biography `bioClean`.

1. 
```{r}
bioClean2 = bioClean %>%
  separate(Hometown, c("City","State"), sep = ", ", remove = TRUE)

head(bioClean2)
```

2. 
```{r}
table(bioClean2$State)
```

## Part three: Joining data frames

1.

```{r}
offbio = left_join(offClean, bioClean, by = "Name")
offbio %>% select("Name", "Hometown") %>% unique()
```

This is a list of all the offensive players and where they are from.

2.
```{r}
df1 = inner_join(offClean, bioClean, by = "Name")
ggplot(df1, aes(x = `Weight`, y = `Receiving_YDS`)) + geom_point() + labs(title = "Relationship between Weight and Receiving Yards", x='Weight', y='Receiving Yards') + geom_smooth()
```

Based on this graph there is no correlation between weight and receiving yards.


3.

Question: Is there a correlation between Height and solo tackles.

```{r}
def_bio = inner_join(defClean, bioClean, by = "Name")
ggplot(def_bio, aes(x = `Height`, y = `Tackles_Solo`)) + geom_point() + geom_smooth()
```

Answer: Based on the scattler plot above, there is no meaningful correlation between number of solo tackles and player Height.
